---
title: "Computational Musicology"
author: "Stijn van der Zalm"
date: "2/12/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(spotifyr)
library(ggplot2)
```

2020 was the year of change for every one of us. Traveling has been brought back to a minimum, schools close and working from home became the standard. This last aspect is interesting to me, since working in another environment comes with other habits. I wonder how the music played working at homes differs from music at the office. 
Luckily Spotify wouldn't be Spotify if it didn't turn every human habit into a playlist. When you search for working from home, you end up with hundreds and hundreds of playlists. I decided to limit this corpus to a special "working from home" folder with all playlists for working from home, created by Spotify. The folder consists of the following playlists:

```{r setup, include=FALSE}
playlist_names <- data.frame("Name" = c("Werkdag FM", "Focus Modus", "To Do List", "Calm & Focused", "Maximum Concentration", "Lo-Fi Beats", 'Lofi Hip-Hop', 'Lush Lofi', 'Instrumental Study', "Jazz Vibes", "Coffee Table Jazz", "Deep Focus"), "Number of songs" = c("250", "59", "100", "102", "180", "450", "100", "190", "285", "300", "143", "194"))
```

```{r}
as_tibble(playlist_names)
```

```{r}
werkdag_fm <- get_playlist_audio_features("", "37i9dQZF1DX4bw5oBAFpuz")
focus_modus <- get_playlist_audio_features("", "37i9dQZF1DWVRrbkzYIlbi")
to_do_list <- get_playlist_audio_features("", "37i9dQZF1DX8sKnLLx9deI")
calm_focused <- get_playlist_audio_features("", "37i9dQZF1DX9etR9S28cIo")
maximum_concentration <- get_playlist_audio_features("", "37i9dQZF1DX0wdQKgYFb7Q")
lofi_beats <- get_playlist_audio_features("", "37i9dQZF1DWWQRwui0ExPn")
lofi_hiphop <- get_playlist_audio_features("", "37i9dQZF1DX36Xw4IJIVKA")
lush_lofi <- get_playlist_audio_features("", "37i9dQZF1DXc8kgYqQLMfH")
instrumental_study <- get_playlist_audio_features("", "37i9dQZF1DX9sIqqvKsjG8")
jazz_vibes <- get_playlist_audio_features("", "37i9dQZF1DX0SM0LYsmbMT")
coffee_table_jazz <- get_playlist_audio_features("", "37i9dQZF1DWVqfgj8NZEp1")
deep_focus <- get_playlist_audio_features("", "37i9dQZF1DWZeKCadgRdKQ")

all_playlists <- bind_rows(
  werkdag_fm %>% mutate(category = "Werkdag FM"), 
  focus_modus %>% mutate(category = "Focus Modus"), 
  to_do_list %>% mutate(category = "To Do List"), 
  calm_focused %>% mutate(category = "Calm & Focused"), 
  maximum_concentration %>% mutate(category = "Maximum Concentration"), 
  lofi_beats %>% mutate(category = "Lofi Beats"), 
  lofi_hiphop %>% mutate(category = "Lofi hip-hop"), 
  lush_lofi %>% mutate(category = "Lush Lofi"), 
  instrumental_study %>% mutate(category = "Instrumental study"), 
  jazz_vibes %>% mutate(category = "Jazz Vibes"), 
  coffee_table_jazz %>% mutate(category = "Coffee Table Jazz"), 
  deep_focus %>% mutate(category = "Deep Focus"))
```

```{r}
ggplot(all_playlists, aes(
  x = energy, 
  y = loudness, 
  color = category
)) + 
  geom_point(size = 0.1)
```









